Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations

### 2024-05-12 index.qmd
chore: update dates and move updates to a different file

# Updates

- [x] Why ARTEMIS electron data and Wind ion data?
- [ ] Disagreement between anisotropy observed and predicted by the theory
  - maybe just too localized anisotropy that is not well measured with available time resolution... 
  - or maybe more complex ion distribution, where nongyrotropy is more important than anisotropy
# TODO

- [ ] Plot Bl(t), jm(l) and ion anisotropy for these cases?

## 02/11/2024 (Encounter 4)

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc4/thickness_current)

![](figures/enc4/density_distribution)

![](figures/enc4/dvl)
:::

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc4/dn-dT)

![](figures/enc4/dT-dB)

![](figures/enc4/dn-dB)
:::

## 02/09/2024 (Encounter 2)

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc2/thickness_current)

![](figures/enc2/density_distribution)

![](figures/enc2/dvl)
:::

Pair plot

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc2/dn-dT)

![](figures/enc2/dT-dB)

![](figures/enc2/dn-dB)
:::

##

![](images/paste-20.png)

Given timeseries $(R_a, \phi_a$, $R_b(t)$,

![](images/paste-11.png)

![](images/paste-12.png)

# Alignment

Composition and ionization state should remain invariant.

PSP_ISOIS-EPILO_L2-IC:   Info  Metadata
Parker Solar Probe ISOIS EPI-Lo Level 2 Ion Composition - David McComas (Princeton University)

>>>>

# Motivation

Studying the radial evolution of solar wind discontinuities from synergistic observations of PSP and Earth-orbiting missions (ARTEMIS, Wind) during aligned intervals.

- How does the discontinuities change with the radial distance from the Sun?

- How is solar wind discontinuities formed? What is the physical mechanisms?

  - Generated at or near the sun?

  - Locally generated in the interplanetary space by turbulence?

# Methods

## In-situ synergistic observations (Alignment)

**Similar** type of solar wind or **Same** solar solar wind

How to define alignment?

- Trajectory & Orbits (theory): ballistic approximation of Parker spiral

- Plasma properties (observation):

  - Invariants: plasma composition, ionization states, mass flux $n u r^2$

  - Correlation: magnetic field / plasma velocity

- Simulation: Trace plasma parcels evolution

<!-- sensing solar wind emitted from the same region on the solar surface, with the caveat that dynamical processes at the chromosphere–corona interface have not dramatically altered the solar footpoint within the time corresponding to the radial propagation from the measurement of one spacecraft to that of the other. -->

### 2024-05-12 presentations/rMeeting.qmd
chore: update dates and move updates to a different file

---
title: "Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
  - name: Chen Shi
  - name: Zesen Huang
# format: revealjs
format:
  html: default
  pptx:
    slide-level: 2
  revealjs:
    output-file: 'sparthb.revealjs.html'
slide-number: true
---

## Solar wind discontinuities (SWD)

Discontinuous changes in plasmas parameters and magnetic fields

- Import element of solar wind turbulences

- Very kinetic structure: interact with energetic part of the solar wind ions

## Formation of SWD

Turbulent ion heating by reconnection of tangential discontinuities. (a) Current density (color) and magnetic field lines, possible reconnection sites are indicated by crosses. (b) The proton temperature anisotropy (color) that indicates acceleration

Alfven wave steepening results in the discontinuity formation: magnetic energy transport from large scale (Alfven waves generated in the solar corona) to small (ion kinetic) scales of discontinuities, where this energy can be dissipated to particle heating

## Motivation

Studying the radial evolution of solar wind discontinuities from synergistic observations of PSP and Earth-orbiting missions (ARTEMIS, Wind) during aligned intervals.

- How does the discontinuities change with the radial distance from the Sun?

- How is solar wind discontinuities formed? What is the physical mechanisms?

  - Generated at or near the sun?

  - Locally generated in the interplanetary space by turbulence?

## In-situ synergistic observations

**Similar** type of solar wind or **Same** solar solar wind

How to define alignment?

- Trajectory & Orbits (theory): ballistic approximation of Parker spiral

- Plasma properties (observation):

  - Invariants: plasma composition, ionization states, mass flux $n u r^2$

  - Correlation: magnetic field / plasma velocity

- Simulation: Trace plasma parcels evolution

<!-- sensing solar wind emitted from the same region on the solar surface, with the caveat that dynamical processes at the chromosphere–corona interface have not dramatically altered the solar footpoint within the time corresponding to the radial propagation from the measurement of one spacecraft to that of the other. -->

## Orbits

:::: {.columns}

::: {.column}
![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E4_orbit.png)
:::

::: {.column}
![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/Lag_from_PSP_to_STA_and_Earth.png)

![Heliographic frame (corotating with Sun)](../notebooks/events/images/paste-1.png)
:::

::::

## ENLIL simulation

::: {.content-visible when-format="html"}

![Normalized solar wind density at the ecliptic and the IMF lines](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_den1-psp-x15wifs3z1fo_med_201903-201904.mp4)

![Radial solar wind velocity at the ecliptic on the global and detail regions together with the temporal profile at spacecraft](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_vel2e1-psp-x15ifs3z1fr_med_201903-201904.mp4)
:::

## Observation

:::: {.columns}

::: {.column}
![PSP Observation Overvall](../images/paste-19.png)
:::

::: {.column}
![PSP observation](../images/paste-18.png)
:::

::::

## Observation (continued)

:::: {.columns}

::: {.column}
![PSP observation](../images/paste-18.png)
:::

::: {.column}
![ACE observation](../images/paste-16.png)
:::

::: {.column}
![ACE observation (continued)](../images/paste-17.png)
:::
::::

Time resolution (from high to low): SPC, SPI, SQTN, QTN

Quality (fluctuating, from high to low): SQTN, QTN, SPC/SPI

![Different Density Product ](../figures/examples/psp_density_comparison.png)

>>>>

---
title: "Current Sheets in the Solar Wind: Parker Solar Probe Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
format:
  html: default
  pptx:
    slide-level: 2
slide-number: true
date: "2024-05"
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms

See [sparthrb](./sparthb.qmd) for alignment and observation details
-->

### 2024-05-12 presentations/sparthb.qmd
chore: update dates and move updates to a different file

---
title: "Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
  - name: Chen Shi
  - name: Zesen Huang
# format: revealjs
format:
  html: default
  pptx:
    slide-level: 2
  revealjs:
    output-file: 'sparthb.revealjs.html'
slide-number: true
---

>>>>

---
title: "Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
  - name: Chen Shi
  - name: Zesen Huang
# format: revealjs
format:
  html: default
  pptx:
    slide-level: 2
  revealjs:
    output-file: 'sparthb.revealjs.html'
slide-number: true
date: "2024-03"
---
